Mental Health Plays Significant Role in Physical Wellness in 1972 to 2022 US General Social Surveys

### 2024-04-17 paper/paper.qmd
update readme and paper

---
title: "Mental Health Plays Significant Role in Physical Wellness in 1972 to 2022 US General Social Surveys"
subtitle: "Mental Illness Being Overlooked in the Medical Field with Gender Disparities"
author: 
  - Mingjia Chen
thanks: "Code and data are available at: https://github.com/MjChen120/Mental_to_Physical_Health.git."
date: today
date-format: long
abstract: "Mental health is a relatively newly introduced topic in the field, but would mental health issues influence physical health as well? Focusing on numerous variables correlated to mental and physical health, this study uses data from the U.S. General Social Survey (GSS) from 1972 to 2022 to investigate the impact of mental health unwellness on physical health unwellness, focusing specifically on “reported days of feeling unwell in a month.” Our analysis reveals that mental unwellness leads to an increase in physical unwellness, and identifies disparities between demographic groups, with women disproportionately affected. These findings underscore the necessity for the importance of promoting mental health and supporting underrepresented groups in the medical field.."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Physical and mental health variables such as "health", "depress" status, and reported days of feeling physical ("physhlth") and health ("mntlhlth") unwell in a month are selected from GSS [@GSS] for main analysis. Variable "year" is chosen for checking data consistency and non-response rate over the years. Demographic data, specifically age and gender, were used as well for further analyses. The data was downloaded and filtered for the selected variables from the selected data variables from GSS[^1]. The data cleaning was performed based on value definitions as defined in the GSS codebooks [@gssCodebook]. The variable names are renamed to be more informative (@tbl-use-var).

@@

@@

## Survey Methodology

@@

# Differences Detected betwen Demographic Populations

>>>>

---
title: "Mental Health Plays Significant Role in Physical Wellness in 1972 to 2022 US General Social Surveys"
subtitle: "Mental Illness Being Overlooked in the Medical Field with Gender Disparities"
author: 
  - Mingjia Chen
thanks: "Code and data are available at: https://github.com/MjChen120/Mental_to_Physical_Health.git."
date: today
date-format: long
abstract: "Mental health is a relatively newly introduced topic in the field, but would mental health issues influence physical health as well? Focusing on numerous variables correlated to mental and physical health, this study uses data from the U.S. General Social Survey (GSS) from 1972 to 2022 to investigate the impact of mental health unwellness on physical health unwellness, focusing specifically on “reported days of feeling unwell in a month.” Our analysis reveals that mental unwellness leads to an increase in physical unwellness, and identifies disparities between demographic groups, with women disproportionately affected. These findings underscore the necessity for the importance of promoting mental health and supporting underrepresented groups in the medical field."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Physical and mental health variables such as "health", "depress" status, and reported days of feeling physical ("physhlth") and health ("mntlhlth") unwell in a month are selected from GSS [@GSS] for main analysis. Variable "year" is chosen for checking data consistency and non-response rate over the years. Demographic data, specifically age and gender, were used as well for further analyses. The data was downloaded and filtered for the selected variables from the selected data variables from GSS[^1]. In total, 72390 responses (with 56% from female respondents) containing mentioned variables were collected. The data cleaning was performed based on value definitions as defined in the GSS codebooks [@gssCodebook]. The variable names are renamed to be more informative (@tbl-use-var).

@@

```{r}
#| label: tbl-sex
#| tbl-cap: Number of Responses from Each Gender Group
#| echo: false
#| warning: false
#| message: false

genderdata <- filter(GSS, sex == 2)

sex <- c("Male","Female")
numbers <- c(nrow(genderdata), nrow(GSS)-nrow(genderdata))
female_per <- round(nrow(genderdata)/nrow(GSS),2) * 100

gendertable <- cbind(sex,numbers)
kable(gendertable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Sex","Responses"))

@@

```{r}
#| label: fig-countedDays
#| tbl-cap: Number of Responses from Each Gender Group
#| echo: false
#| warning: false
#| message: false

```
## Survey Methodology

@@

## Differences Detected betwen Demographic Populations
